\textbf{The Mint Scale: A Fresh Validation of the Multimodal Interoception Questionnaire and Comparison to the MAIA, BPQ and IAS}

### 2025-07-02 paper/manuscript.tex
update

\authorsnames[{1,2},{1},{1}]{Dominique Makowski,Ana Neves,Giulia
Poreiro}

@@

\leftheader{Makowski, Neves and Poreiro}

@@

\authornote{\par{\addORCIDlink{Dominique
Makowski}{0000-0001-5375-9967}}\par{\addORCIDlink{Ana
Neves}{0009-0006-0020-7599}}\par{\addORCIDlink{Giulia
Poreiro}{0000-0002-2343-5109}} 
\par{ }
\par{     \begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, toprule=.15mm, arc=.35mm, left=2mm, breakable, colback=white, rightrule=.15mm, bottomrule=.15mm, leftrule=.75mm, opacityback=0]

@@

\end{tcolorbox}  Author roles were classified using the Contributor Role
Taxonomy (CRediT; https://credit.niso.org/) as follows:  Dominique
Makowski:   Conceptualization, Data curation, Formal Analysis, Funding
acquisition, Investigation, Methodology, Project
administration, Resources, Software, Supervision, Validation, Visualization, Writing
-- original draft; Ana Neves:   Investigation, Methodology, Project
administration, Supervision, Writing -- original draft, Writing --
review \& editing; Giulia Poreiro:   Investigation, Methodology, Writing
-- original draft, Writing -- review \& editing}
\par{Correspondence concerning this article should be addressed
to Dominique
Makowski, Email: \href{mailto:D.Makowski@sussex.ac.uk}{D.Makowski@sussex.ac.uk}}
}

@@

TODO: write general intro.

Main issues in existing questionnaires: - Either heavily based on
theories (e.g., focusing on a specific dimensions), despite shaky
evidence for said-theories - Do not control for context (which leads to
variability in interpretation and occurence) - Often quite narrow in the
modalities covered

@@

Goal of study 1: to generate a lot of items, analyze its structure and
reduce them to a balanced set of items.

@@

We recruited 760 English-speaking participants using
Prolific\textcopyright. We excluded 191 for failing at least one
attention check, and 10 based on measures significantly related to the
probability of failing attention checks (namely, the multivariate
distance obtained with the OPTICS algorithm,
\citeproc{ref-theriault2024check}{Thériault et al., 2024}). The final
sample includes 559 participants (age = 37.0 \(\pm\) 12.2 {[}18, 77{]};
50.8\% women; Country of residence: 63.86\% UK, 26.65\% USA). This study
was approved by the University of Sussex' Ethics Committee
(\textbf{NUMBER}).

@@

Based on the two goals outlined for this scale, namely to include
different interoceptive modalities, and to explicitly state the context
of the interoceptive experience (e.g., whether negative or positive), we
generated items in a systematic way following a combinatorial approach,
where each item's category was a combination of a specific modality and
context (Figure~\ref{fig-one}).

@@

{\caption{{The conceptual grid used to generate the 120 initial items
(top-left). Each item belong both to an interoceptive modalitiy and a
facet, with the number of each item per category indicated in the
circles. The askterisk denotes the additional presence of an attention
check item in that category. In the experiment, these items were
presented on different pages grouped either by modality (bottom-left),
by facet (top-right), or entirely randomly. The Correlation Similarity
(bottom-right) analysis suggested that the correlation matrix obtained
from the participants assigned to the random-grouping condition was
slightly more similar (but non-significantly) to the one obtained in the
modality-grouping condition, suggesting that 1) the scale's structure is
robust to different presentation conditions; 2) modality-grouping might
tend to facilitate the emergence of the underlying item structure (and
thus be interpreted as being more natural).}{\label{fig-one}}}}

@@

We firstly identified 7 ``modalities'' (cardiac, respiratory, gastric,
genital, skin \& temperature, bladder \& colon, and a ``general state''
category corresponding to a holistic and general awareness of an
interoceptive state or dimension). Through iterative refinement (e.g.,
splitting or merging different categories together), we then settled on
6 ``facets'', which encompass both \emph{contexts} of experience
(negative and positive arousal, namely anxious and sexual states), and
potential distinct \emph{mechanisms} (nociception \& pleasure,
sensitivity, accuracy, and confusion).

Using this orthogonal 7x6 modality/facet grid as a conceptual
scaffolding, we generated 120 initial items, striving for a balanced
number of items with consistent phrasing within modalities and
facets\footnote{The initial item list at
  \href{https://realitybending.github.io/InteroceptionScale/study1/analysis/2_analysis.html}{realitybending.github.io/InteroceptionScale/study1/analysis/2\_analysis.html}}.
We additionally crafted 8 ``attention check'' items blending in (and
distributed across) each category.

@@

To avoid presenting all the 120 items on a single long and discouraging
page, we split them into different pages. Participants were randomly
assigned to one of three conditions, driving how items were grouped on
the same page: 1) items grouped by modality (i.e., all cardiac items on
the first page, all colon \& bladder items on the second, etc.), 2)
items grouped by facet, or 3) items presented fully randomly (but
balanced randomly across 6 pages). The order of the item on any given
page and the order of the modalities/facets were randomized. Each
participant completed the full set of 120 items, with the attention
check items interspersed throughout. The online experiment was
implemented using JsPsych (\citeproc{ref-de2015jspsych}{De Leeuw,
2015}), and item responses were recorded using 7-points Likert scales (0
= Disagree, 6 = Agree).

@@

Items deemed ``redundant'' (which can distort the item structure
estimation by introducing multicollinearity or local dependencies) will
be identified (using the recommended threshold of 0.25) using Unique
Variable Analysis (UVA, \citeproc{ref-christensen2023unique}{Christensen
et al., 2023}), a novel and principled method derived from network
psychometrics.

The structure of the items will be analyzed using the recently-developed
Exploratory Graph Analysis (EGA, \citeproc{ref-golino2017exploratory}{H.
F. Golino \& Epskamp, 2017}) framework, which allows to jointly estimate
the number of dimensions (i.e., clusters of items), the structure, as
well as its stability using bootstrapping
(\citeproc{ref-golino2020investigating}{H. Golino et al., 2020}). At a
fundamental level, EGA conceptualizes variables as nodes in a network,
with connections (edges) reflecting associations between them. Evidence
has underlined its suitability as an alternative to traditional factor
analysis, addressing some of its limitations such as the assumption of a
``latent'' source of variability, issues in estimation of the optimal
factor numbers, and poor performance in complex population structures,
while remaining comparable and interpretable
(\citeproc{ref-christensen2021equivalency}{Christensen \& Golino, 2021};
\citeproc{ref-jimenez2023dimensionality}{Jiménez et al., 2023}). In
particular, nodes communities (i.e., clusters of items) can be in
practice interpreted as distinct ``dimensions'', similarly to
traditional latent factors - but without explicitly assuming their
existence (\citeproc{ref-christensen2021equivalency}{Christensen \&
Golino, 2021}).

After removing redundant items using UVA, we will iteratively fit
hierarchical EGA models (which additionally estimates higher-order
``meta'' clusters) using ``glasso'' {[}\textbf{REF}{]} and the
``leiden'' algorithm {[}\textbf{REF}{]} for community detection,
refining the item pool at each step. We will start by removing items
with a low (\textless{} 80\%) cluster stability (i.e., volatile items
which jump between clusters across bootstrapped samples), followed by
odd items belonging to no clusters or pairs of items (i.e., we keep
items belonging to clusters of more than 2 items). Finally, for each
lower-level cluster, we will select the 3 items with the highest node
centrality (i.e., the highest loading in the cluster).

@@

The correlation matrix similarity analysis yielded no significant
differences between the similarity of the random-grouping condition with
the modality-grouping and facet-grouping conditions
(\(PSI_{\text{Random vs. Facet}} = 0.81\),
\(PSI_{\text{Random vs. Modality}} = 0.82\), \(p = .45\);
\(RVadj_{\text{Random vs. Facet}} = 0.77\),
\(RVadj_{\text{Random vs. Modality}} = 0.78\), \(p = .74\);
\(SMI_{\text{Random vs. Facet}} = 0.49\),
\(SMI_{\text{Random vs. Modality}} = 0.51\), \(p = .52\)).

From the 120 initial items, UVA flagged 4 redundant items that we
removed. We then removed 40 items that showed low cluster stability, and
9 items that were part of clusters with less than 3 items. Finally, We
kept the 3 items with the highest loading in their lower-level structure
(removing 13 items in the process), resulting in 54 items in the final
item pool.

@@

Possible (although not at all significant) bias consistent across
indices in favour of a greater similarity with the modality-grouping
condition.

@@

TODO: write intro. Goal of study 2: to validate the Mint scale against
other interoception scales.

@@

We recruited 921 English-speaking participants via
Prolific\textcopyright and SONA, from which 118 were excluded for
failing at least one attention check and 6 based on multivariate
distance (using the same procedure as for study 1). 60 participants were
further excluded due to missing data following a technical error. The
final sample includes 737 participants (age = 36.8 \(\pm\) 14.7 {[}18,
87{]}; 57.3\% women; Country of residence: 75.17\% United Kingdom,
24.83\% other). This study was approved by the University of Sussex'
Ethics Committee (\textbf{NUMBER}).

The experiment included the following questionnaires.

@@

\subparagraph{The Multimodal Interoception Questionnaire
(Mint).}\label{the-multimodal-interoception-questionnaire-mint}

The 45 items of the Mint scale were presented in a random order, with
the same 7-point Likert scale as in study 1 (0 = Disagree, 6 = Agree).

\subparagraph{Multidimensional Assessment of Interoceptive Awareness
(MAIA).}\label{multidimensional-assessment-of-interoceptive-awareness-maia}

The X items of the MAIA\ldots{}

\subparagraph{The Body Perception Questionnaire
(BPQ).}\label{the-body-perception-questionnaire-bpq}

\textbf{TODO}

\subparagraph{The Interoceptive Awareness Scale
(IAS).}\label{the-interoceptive-awareness-scale-ias}

\textbf{TODO}

@@

\subparagraph{Toronto Alexithymia Scale
(TAS-20).}\label{toronto-alexithymia-scale-tas-20}

\textbf{TODO}

\subparagraph{Emotion Reactivity Scale
(ERS).}\label{emotion-reactivity-scale-ers}

\textbf{TODO}

\subparagraph{Cognitive Emotion Regulation Questionnaire
(CERQ).}\label{cognitive-emotion-regulation-questionnaire-cerq}

\textbf{TODO}

\subparagraph{TODO (CEFSA).}\label{todo-cefsa}

\textbf{TODO}

\subparagraph{Primal World Beliefs
(PI-18).}\label{primal-world-beliefs-pi-18}

\textbf{TODO}

@@

\subparagraph{Life Satisfaction, Depression and Anxiety
(PHQ-4).}\label{life-satisfaction-depression-and-anxiety-phq-4}

\textbf{TODO}

\subparagraph{Mental Health.}\label{mental-health}

@@

\subparagraph{Somatic Health.}\label{somatic-health}

@@

\textbf{TODO}

Physical activity.

Wearables.

\paragraph{Randomization.}\label{randomization}

In order to avoid the repetition of similar types of questions and and
balance longer and shorter questionnaires, we partitioned them into
three groups (and randomized their order within them): (1) interoception
questionnaires (MAIA, IAS, BPQ), (2) emotions (TAS, CERQ, ERS, PI-18),
and (3) pathology (somatic health, mental health, LS + PHQ-4, CEFSA).
After completing demographic questions, participants always started with
the Mint scale, and each following interoception questionnaire was
interspersed with two questionnaires from the emotions and pathology
groups. In order to make the experiment more enjoyable, the experiment
ended with a radar chart summarizing the participants' responses to the
Mint scale\footnote{The experiment can be tested by following the link
  on https://github.com/RealityBending/InteroceptionScale}.

We will start by confirming and further refining the structure of the
Mint scale using the same EGA model as in Study 1. The convergent
validity of the final set of items will be assessed by computing the
correlations between the Mint scale and the other interoception
questionnaires (MAIA, BPQ, IAS). Then, we will test the predictive power
of the Mint scale (relative to the other interoception questionnaires)
on various measures. We will start by fitting regression (linear for
continuous measures - e.g., depression score - and logistic for binary

@@

{\caption{{TODO.}{\label{fig-two}}}}

\begin{center}
\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{../study2/analysis/figures/fig1.png}
\end{center}

{\caption{{TODO.}{\label{fig-three}}}}

@@

@@

{\caption{{TODO.}{\label{fig-four}}}}

@@

>>>>

\authorsnames[{1,2},{1},{1},{1},{1}]{Dominique Makowski,Ana Neves,Emma
Benn,Maisie Bennett,Giulia Poreiro}

@@

\leftheader{Makowski, Neves, Benn, Bennett and Poreiro}

@@

\authornote{\par{\addORCIDlink{Dominique
Makowski}{0000-0001-5375-9967}}\par{\addORCIDlink{Ana
Neves}{0009-0006-0020-7599}}\par{\addORCIDlink{Giulia
Poreiro}{0000-0002-2343-5109}} 
\par{ }
\par{     \begin{tcolorbox}[enhanced jigsaw, toprule=.15mm, colframe=quarto-callout-note-color-frame, colback=white, opacityback=0, left=2mm, arc=.35mm, leftrule=.75mm, breakable, bottomrule=.15mm, rightrule=.15mm]

@@

\end{tcolorbox}  Author roles were classified using the Contributor Role
Taxonomy (CRediT; https://credit.niso.org/) as follows:  Dominique
Makowski:   Conceptualization, Data curation, Formal Analysis, Funding
acquisition, Investigation, Methodology, Project
administration, Resources, Software, Supervision, Validation, Visualization, Writing
-- original draft; Ana Neves:   Investigation, Methodology, Project
administration, Supervision, Writing -- original draft, Writing --
review \& editing; Emma Benn:   Investigation; Maisie
Bennett:   Investigation; Giulia
Poreiro:   Investigation, Methodology, Writing -- original
draft, Writing -- review \& editing}
\par{Correspondence concerning this article should be addressed
to Dominique
Makowski, Email: \href{mailto:D.Makowski@sussex.ac.uk}{D.Makowski@sussex.ac.uk}}
}

@@

Main issues in existing questionnaires:

\begin{itemize}
\tightlist
\item
  Either heavily based on theories (e.g., focusing on specific
  ``alleged'' dimensions), despite shaky evidence for said-theories
\item
  Do not control for context (which leads to variability in
  interpretation and occurence)
\item
  Often quite narrow in the modalities covered (or scattered, like the
  IAS)
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  Goal of study 1: to generate a lot of items, analyze its structure and
  reduce them to a balanced set of items.
\end{itemize}

@@

We recruited 760 English-speaking participants using
Prolific\textcopyright. We excluded 191 for failing at least one
attention check, and 10 based on measures significantly related to the
probability of failing attention checks (namely, the multivariate
distance obtained with the OPTICS algorithm,
\citeproc{ref-theriault2024check}{Thériault et al., 2024}). The final
sample includes 559 participants (age = 37.0 \(\pm\) 12.2 {[}18, 77{]};
50.8\% women; country of residence: 63.86\% UK, 26.65\% USA). This study
was approved by the University of Sussex' Ethics Committee
(ER/MB2021/1).

@@

Based on the two goals outlined for this scale, namely to include
different interoceptive modalities, and to explicitly state the context
of the interoceptive experience (e.g., whether negative or positive), we
generated items in a systematic way following a combinatory approach,
where each item's category corresponds to the union of a specific
modality and context (Figure~\ref{fig-one}).

@@

{\caption{{The conceptual grid used to generate the 120 initial items
(top-left). Each item belongs both to an interoceptive modality
(vertical) and a facet (horizontal), with the number of each item per
category indicated in the circles. The asterisk denotes the additional
presence of an attention check item in that category. In the experiment,
these items were presented on different pages grouped either by modality
(bottom-left), by facet (top-right), or randomly. The Correlation
Similarity (bottom-right) analysis suggested that the correlation matrix
obtained from the participants assigned to the random-grouping condition
was slightly more similar (but non-significantly) to the one obtained in
the modality-grouping condition, suggesting that the scale's structure
is robust to different presentation conditions, and that the
modality-grouping might potentially tend to facilitate the emergence of
the underlying item structure (and thus be interpreted as being more
``natural'').}{\label{fig-one}}}}

@@

We identified 7 ``modalities'' (cardiac, respiratory, gastric, genital,
skin \& temperature (thermoregulation), bladder \& colon, and a
``general state'' category corresponding to a holistic and general
awareness of an interoceptive state or dimension). Through iterative
refinement (e.g., splitting or merging different categories together),
we settled on 6 ``facets'', which encompass both \emph{contexts} of
experience (negative and positive arousal, e.g., anxious and sexual
states), and potential distinct \emph{mechanisms} (nociception \&
pleasure, sensitivity, accuracy, and confusion).

Using this orthogonal 7x6 modality/facet grid as a conceptual
scaffolding, we generated 120 initial items, striving for a balanced
number of items with consistent phrasing within modalities and
facets\footnote{The initial item list is available at
  \href{https://realitybending.github.io/InteroceptionScale/study1/analysis/2_analysis.html}{realitybending.github.io/InteroceptionScale/study1/analysis/2\_analysis.html}}.
We additionally crafted 8 ``attention check'' items blending in (and
distributed across) each category.

@@

To avoid presenting all the 120 items on a single long and discouraging
page, we split them into different pages. Participants were randomly
assigned to one of three conditions, driving how items were grouped on
the same page: 1) items grouped by modality (i.e., all cardiac items on
the first page, all colon \& bladder items on the second, etc.), 2)
items grouped by facet, or 3) items presented fully randomly (but with
their number balanced across 6 pages). The order of the item on any
given page and the order of the modalities/facets was randomized. Each
participant completed the full set of 120 items, with 8 attention check
items interspersed throughout. The online experiment was implemented
using JsPsych (\citeproc{ref-de2015jspsych}{De Leeuw, 2015}), and item
responses were recorded using 7-points Likert scales (0 = Disagree, 6 =
Agree).

@@

Items deemed ``redundant'' (which can distort the item structure
estimation by introducing multicollinearity or local dependencies) were
identified (using the recommended threshold of 0.25) using Unique
Variable Analysis (UVA, \citeproc{ref-christensen2023unique}{Christensen
et al., 2023}), a novel and principled method derived from network
psychometrics.

The structure of the items was analyzed using the recently-developed
Exploratory Graph Analysis (EGA, \citeproc{ref-golino2017exploratory}{H.
F. Golino \& Epskamp, 2017}) framework, which allows to jointly estimate
the number of dimensions (i.e., clusters of items), the structure, as
well as its stability using bootstrapping
(\citeproc{ref-golino2020investigating}{H. Golino et al., 2020}). At a
fundamental level, EGA conceptualizes variables as nodes in a network,
with connections (edges) reflecting associations between them. Evidence
has underlined its suitability as an alternative to traditional factor
analysis, addressing some of its limitations such as the assumption of a
``latent'' source of variability, issues in estimation of the optimal
factor numbers, and poor performance in complex population structures,
while remaining comparable and interpretable
(\citeproc{ref-christensen2021equivalency}{Christensen \& Golino, 2021};
\citeproc{ref-jimenez2023dimensionality}{Jiménez et al., 2023}). In
particular, nodes communities (i.e., clusters of items) can be in
practice interpreted as distinct ``dimensions'', similarly to
traditional latent factors - but without explicitly assuming their
existence (\citeproc{ref-christensen2021equivalency}{Christensen \&
Golino, 2021}).

After removing redundant items using UVA, we iteratively fitted
hierarchical EGA models (which additionally estimates higher-order
``meta'' clusters) using ``glasso''
(\citeproc{ref-friedman2008sparse}{Friedman et al., 2008}) and the
``Leiden'' algorithm (\citeproc{ref-traag2019louvain}{Traag et al.,
2019}) for community detection, refining the item pool at each step. We
started by removing items with a low (\textless{} 80\%) cluster
stability (i.e., volatile items which jump between clusters across
bootstrapped samples), followed by odd items belonging to no clusters or
pairs of items (i.e., we keep items belonging to clusters of more than 2
items). Finally, for each lower-level cluster, we selected the 3 items
with the highest node centrality (i.e., the highest loading in the
cluster).

@@

The correlation matrix similarity analysis yielded no significant
differences between the similarity of the random-grouping condition with
the modality-grouping and facet-grouping conditions
(\(PSI_{\text{Random vs. Facet}} = 0.81\),
\(PSI_{\text{Random vs. Modality}} = 0.82\), \(p = .45\);
\(RVadj_{\text{Random vs. Facet}} = 0.77\),
\(RVadj_{\text{Random vs. Modality}} = 0.78\), \(p = .74\);
\(SMI_{\text{Random vs. Facet}} = 0.49\),
\(SMI_{\text{Random vs. Modality}} = 0.51\), \(p = .52\)), despite a
consistent bias in favour of the modality-grouping condition.

From the 120 initial items, 4 redundant items were flagged by UVA. We
then removed 40 items that showed low cluster stability, and 9 items
that were part of clusters with less than 3 items. Finally, We kept the
3 items with the highest loading in their lower-level structure
(removing 13 items in the process), resulting in 54 items in the final
item pool.

@@

\begin{itemize}
\tightlist
\item
  Possible (although not at all significant) bias consistent across
  indices in favour of a greater similarity with the modality-grouping
  condition. Futher research with more data (here we only had
  \textasciitilde200 observation per condition) might be able to yield
  more conclusive results. In general, testing the presentation mode is
  a novel approach which should be more widely applied to psychometric
  validations.
\item
  We grouped some of the dimensions under ``Awareness'', despite being
  similar to items included in the IAS typically conceptualized as
  ``Accuracy''. However, we believe that subjective scales cannot be
  pure measures of accuracy, as they are inherently subjective and thus
  reflect a degree of awareness (of one's accurate noticing or
  sensation).
\item
  Olfactory compensation: pseudo-interoceptive pathway whereby people
  with interoceptive deficits would compensate by gathering evidence
  about their internal state externally (e.g., via smell).
\item
  The metacluster ``Sensitivity'' seems to be grouping iterms related to
  visceropcetion and (hyper)sensitivity (Derm).
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  Goal of study 2: to validate the Mint scale against other
  interoception scales.
\end{itemize}

@@

We recruited 921 English-speaking participants via SONA and
Prolific\textcopyright, from which 118 were excluded for failing at
least one attention check and 6 based on multivariate distance (using
the same procedure as for study 1). 60 participants were further
excluded due to missing data following a technical issue. The final
sample includes 737 participants (age = 36.8 \(\pm\) 14.7 {[}18, 87{]};
57.3\% women; Country of residence: 75.17\% United Kingdom, 24.83\%
other). This study was approved by the University of Sussex' Ethics
Committee (ER/EB672/2).

\subsubsection{Measures}\label{measures}

{\caption{{Structure analysis of the final set of 33 Mint items.
Hierarchical Exploratory Graph Analysis (EGA) was applied to jointly
identify clusters of items, as well as higher-order metaclusters. The
reliability of each item can be quantified by the proportion of
structure replication across bootstrapped samples, with high values
indicative of high structural stability. The Mint scale displayed an
excellent structural consistency, with the exception of the \emph{Olfa}
cluster which belonging to the \emph{Deficit} metaclusters warrants
further research. Bottom plot shows the result of hierarchical
clustering, providing evidence for structure statbility across methods,
and allowing for more granular understanding of the relationships
between variables.}{\label{fig-two}}}}

\begin{center}
\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{../study2/analysis/figures/fig1.png}
\end{center}

@@

The 45 items of the Multimodal Interoception Questionnaire (Mint) scale
were presented in a random order, with the same 7-point Likert scale as
in study 1 (0 = \emph{Disagree}, 6 = \emph{Agree}).

The Multidimensional Assessment of Interoceptive Awareness (MAIA-2,
\citeproc{ref-mehling2018multidimensional}{Mehling et al., 2018})
measures 8 dimensions with 37 items presented on a 7-point Likert scale
(0 = \emph{Never}, 6 = \emph{Always}). It includes bodily dimensions
such as Body Listening, Noticing, Emotional Awareness, dimensions
related to emotion regulation, such as Trust, Not-worrying, and
Self-Regulation, as well as dimensions related to attention, such as
Attention Regulation and Not-Distracting. The relationship of some of
these dimensions to interoception remains debated, particularly
Not-Distracting and Not-Worrying, which tend to show weak or
non-existent correlations with the other MAIA subscales
(\citeproc{ref-ferentzi2021examining}{Ferentzi et al., 2021}).

The Body Perception Questionnaire - Very Short Form (BPQ-VSF,
\citeproc{ref-cabrera2018assessing}{Cabrera et al., 2018}) measures a
general score of bodily awareness with 12 items presented on 5-point
Likert scale (0 = \emph{Never}, 5 = \emph{Always}).

The Interoceptive Accuracy Scale (IAS,
\citeproc{ref-murphy2020testing}{Murphy et al., 2020}) measures a
general score of interoceptive accuracy to physical sensations with 21
items presented on a 5-point Likert scale (1 = \emph{Disagree Strongly},
5 = \emph{Strongly Agree}).

@@

The Toronto Alexithymia Scale TAS (TAS-20,
\citeproc{ref-leising2009toronto}{Leising et al., 2009}) measures 3
Alexithymia dimensions with 20 items 5-point Likert scale (1 =
\emph{Strongly Disagree}, 5 = \emph{Strongly Agree}: Difficulty
Identifying Feelings (DIF), Difficulty Describing Feelings (DDF), and
Externally Oriented Thinking (EOR).

The Emotion Reactivity Scale - Brief Version (B-ERS,
\citeproc{ref-veilleux2024development}{Veilleux et al., 2024}) measures
3 dimensions with 6 items on a 5-point Likert scale (0 = \emph{Not like
me at all}, 4 = \emph{Extremely like me}): Arousal (\emph{``My moods are
very strong and powerful''}), Sensitivity (\emph{``I tend to get very
emotional very easily''}), and Persistence (\emph{``When I am
angry/upset, it takes me much longer than most people to calm down''}).

The Cognitive Emotion Regulation Questionnaire (CERQ-short,
\citeproc{ref-garnefski2006cognitive}{Garnefski \& Kraaij, 2006})
measures 9 adaptive and maladaptive emotion regulation strategies with
18 items (1 = \emph{Almost Never}, 5 = \emph{Almost Always}). The
strategies include Self-blame, Other-blame, Rumination, Catastrophizing,
Putting into Perspective, Positive Refocusing, Positive Reappraisal,
Acceptance, and Planning.

The Černis Felt Sense of Anomaly scale - short form (CEFSA-14,
\citeproc{ref-vcernis2024measuring}{Černis et al., 2024}) measures 7
dissociative experiences with 14 items on a 5-point Likert scale (0 =
\emph{Never}, 4 = \emph{Always}). This scale is designed to measure
dissociative experiences in adolescence and adulthood, specifically
focusing on the felt sense of anomaly-type dissociation. It includes
Anomalous Experience of the Self, Anomalous Experience of the Body,
Anomalous Experiences of Emotion, Altered Sense of Familiarity, Altered
Sense of Connection, Altered Sense of Agency, and Altered Sense of
Reality.

The Primal World Beliefs Inventory (PI-18,
\citeproc{ref-clifton2021brief}{Clifton \& Yaden, 2021}) measures
higher-order beliefs about the basic character of the world: Good, Safe,
Enticing, and Alive. The 18 items are presented on a 6-point Likert
scale (0 = \emph{Strongly Disagree}, 5 = \emph{Strongly Agree}).

@@

The Single Item Life Satisfaction scale (SILS,
\citeproc{ref-jovanovic2020longer}{Jovanović \& Lazić, 2020}) was
followed by the PHQ-4 (\citeproc{ref-kroenke2009ultra}{Kroenke et al.,
2009}) assessing depression and anxiety symptoms with 4 items. We used
the refined version of the PHQ-4
(\citeproc{ref-makowski2025measuring}{Makowski et al., 2025}), which
adds an additional response option (``Once or twice''), increasing
sensitivity to subclinical fluctuations.

@@

@@

Participants reported about owning any wearable devices to monitor
health indices such as heart rate, number of steps, calories burned,
sleep quality, and weight. For each selected device, a follow-up
question inquired about the frequency of usage and subjective importance
of that measure.

Participants were asked to rate how physically active they consider
themselves and how many hours of active workout they engage in per week.
Participant's BMI was computed using height and weight.

In order to avoid the repetition of similar types of questions and and
balance longer and shorter questionnaires, we partitioned the measures
into three groups (and randomized the order within them): 1)
interoception questionnaires (MAIA, IAS, BPQ), 2) emotions (TAS, CERQ,
ERS, PI-18), and 3) health (somatic symptoms, mental health, LS + PHQ-4,
CEFSA). After completing demographic questions, participants always
started with the Mint scale, and each following interoception
questionnaire was interspersed with two questionnaires from the emotions
and pathology groups. 7 attention checks were embedded throughout (one
within each major questionnaire). In order to make the experiment more
enjoyable, the experiment ended with a radar chart summarizing the
participants' responses to the Mint scale\footnote{The experiment can be
  tested by following the link on
  https://github.com/RealityBending/InteroceptionScale}.

@@

{\caption{{Item loadings. The table shows each item of the Mint with its
cluster centrality, which is the EGA equivalent (in terms of
interpretation) of factor loadings. It also shows how each lower-level
cluster is related to higher-order metaclusters.}{\label{fig-three}}}}

@@

We started by confirming and further refining the structure of the Mint
scale (see Figure~\ref{fig-two}) using the same EGA model as in Study 1.
We then computed the lower-level dimensions and the higher-level
metaclusters' scores by averaging their corresponding items. The
convergent validity of the final set of items was assessed by computing
the correlations between the Mint scale and the other interoception
questionnaires (MAIA, BPQ, IAS).

Next, we tested the predictive power of the Mint scale relative to other
interoception questionnaires. We will assessed the importance of each
interoceptive dimension (from all the scales) as a unique predictor by
fitting regression models (linear for continuous measures - e.g.,
depression score - and logistic for binary variables - e.g., presence
\emph{vs.} absence of mood disorders) to predict different outcomes, and
compare the best between the 4 interoception scales (based on the R2).
We then evaluated the predictive performance of each scale as a whole by
comparing regression models with all the dimensions entered as
predictors (note that the IAS and BPQ only have one total score
variable). We assessed the models based on their total explained
variance (R2), as well as on the Bayesian Information Criterion (BIC),
which penalizes for predictor number, thus offering a balance between
model parsimony and predictive power. For the logistic models, we
quantified the discriminative power by computing the Area Under the
Curve (AUC) of the Receiver Operating Characteristic (ROC) curve, which
assesses the model's discriminative power (the combination of
sensitivity and specificity).

In order to evaluate the potential for a short version of our scale, we
compared 4 variants of the Mint: the (full) \emph{Mint} (including all
lower-level dimensions), the \emph{metaMint} (only including the
metaclusters), the \emph{miniMint} (including the metaclusters computed
from a reduced number of items), and the \emph{microMint} (including
only the most representative dimension of each metacluster). Moreover,
we also included an alternative ``interoception-focused'' version of the
MAIA (\emph{iMAIA}) that only contains the 3 most interoceptive
dimensions (Body Listening, Noticing, and Emotional
Awareness)\footnote{See correlation results to further justify this
  selection.}.

The application of EGA to the initial set of 42 items reproduced the
expected 14 lower-level clusters of triplets and 3 higher-order
metaclusters, with the exception of the \emph{UrSe} (Urointestinal
Sensitivity), for which one item moved to the \emph{Olfa} (Olfactory
Compensation) cluster. In order to further balance and reduce the items,
we removed the \emph{UrSe} cluster, as well as \emph{StaS} (State
Specificity), which in comparison to the other items stood out as
containing vague and overlapping items ; \emph{SexO} (Sexual Organs
Sensitivity), which overlapped with the \emph{SexS} (Sexual Arousal
Sensitivity) cluster; and \emph{CaNo} (Cardiorespiratory Noticing),
which overlapped with the \emph{CaCo} (Cardiorespiratory Confusion)
cluster.

The final set of 33 items yielded a good fit (Generalized Total Entropy
Fit Index = -77.23), with all items showing high cluster stability
(\textgreater{} 90\%), with the exception of \emph{Olfa}. The final
structure included 11 lower-level clusters, grouped into 3 higher-order
metaclusters: ``Interoceptive Deficit'' (containing \emph{CaCo},
\emph{UrIn}, \emph{Derm}, \emph{Sati}, \emph{Olfa}), ``Interoceptive
Awareness'' (containing \emph{ExAc}, \emph{RelA}, \emph{SexS}), and
``Visceroception'' (containing \emph{Card}, \emph{Resp}, \emph{Gast}).
Item centrality (intepretable as loadings in traditional factor analysis
frameworks) are shown in Figure~\ref{fig-three}. Additionally, we
applied hierarchical clustering analysis which replicated this structure
(see Figure~\ref{fig-two}), suggesting consistency across methods.

@@

{\caption{{Correlation Matrices between the Mint dimensions
(upper-left), and between the Mint and other interoception
questionnaires (MAIA, BPQ and IAS; upper-right). The bottom matrix shows
the relationships between interoception and other measures included in
the study, such as alexithymia (TAS-20), depression and anxiety (PHQ-4),
emotion reactivity (ERS), abnormal and dissociative experiences (CEFSA),
and Primal World Beliefs. Stars indicate dimensions that have been
score-reversed for better in-context interpretation. Correlation
coefficients are shown for significant
correlations.}{\label{fig-four}}}}

@@

The correlation matrix of the Mint dimensions revealed an interesting
and rich tapestry of relationships (see Figure~\ref{fig-four}), with
contrasting patterns of associations with dimensions from the same
group. For instance, \emph{Derm} and \emph{Olfa}, despite being
positively correlated with the other dimensions from the \emph{Deficit}
cluster, were not negatively correlated to \emph{Awareness} and its
dimensions. Similarly, \emph{Sati}, unlike the other \emph{Deficit}
dimensions, was not positively correlated to \emph{Visceroception}. This
complex structure suggests that the dimensions indeed capture distinct
aspects of interoception, and that the metaclusters, rather than being
simple aggregates of rather-redundant elements, might actually capture
unique combinations (greater than the sum of their parts). It also
provides evidence against a simplistic adaptive \emph{vs.} maladaptive
dichotomy, as \emph{Deficit} was not negatively bit positively
correlated with \emph{Visceroception}. Interestingly, while
\emph{Awareness} was also positively correlated with
\emph{Visceroception}, it yielded an insignificant negative correlation
with \emph{Deficit} (this finding was also aligned with the hierarchical
clustering results, showing a greater proximity between \emph{Deficit}
and \emph{Visceroception} than with \emph{Awareness}), underlining again
the complex web of relationships captured by the Mint.

The correlation matrix between the Mint and the other interoception
scales revealed high levels of overlap, as well as some unique
contributions (see Figure~\ref{fig-four}). The BPQ was positively
correlated with most Mint dimensions (the highest with
\emph{Visceroception}, \(r = .46\)). The IAS was positively correlated
with \emph{Visceroception} and \emph{Awareness} dimensions (the highest
with \emph{Awareness}, \(r = .63\)), but negatively with most
\emph{Deficit} dimensions (with the exception of \emph{Olfa} and
\emph{Derm}). In turn, the \emph{Visceroception} metacluster most
strongly correlated with MAIA's Noticing (\(r = .55\)). Interestingly,
MAIA's TTrusting correlated selectively with \emph{Awareness}, and
negatively with \emph{Deficit} dimensions, but not with
\emph{Visceroceptive} dimensions (underlining its high-level
metacognitive nature). MAIA's Emotional Awareness and Body Listening
displayed a similar pattern to Noticing, and Attention Regulation and
Self Regulation positively correlated with \emph{Awareness} and
\emph{Visceroception} dimensions, but negatively with some
\emph{Deficit} dimensions (\emph{CaCo} and \emph{UrIn}). Not-distracting
only yielded mild negative associations with \emph{Sati} and
\emph{Olfa}. Overall, the Mint dimensions were able to capture most of
the (relevant) variance and intricacies present in the other
interoception scales.
